Rough calculations

### 2015-10-31 notes/diffusion-local-times.Rmd
rough calcs

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\grad}{\nabla}
$$

@@

>>>>

$$
\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\calE}{\mathcal{E}}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\grad}{\nabla}
$$

@@

Notes on excursion theory
=========================

Let $n(e)$ denote the excursion measure of Brownian motion,
on $\calE = \{ f : [0,\infty) \to [0,\infty) \, \text{s.t.} f(s)=0, t>s \Rightarrow f(t)=0 \}$.
For $f \in \calE$ denote $\zeta(f) = \sup\{ t : f(t)>0\}$.
Then [Rogers 1989](http://www.bibsonomy.org/bibtex/201686d46785a10af443e4d48a5196088/peter.ralph)
$$\begin{aligned}
n(f : \sup_t f(t) > a ) = \frac{1}{2a} .
\end{aligned}$$
and
$$\begin{aligned}
n(f : \zeta f > t ) = \sqrt{\frac{2}{\pi t}} .
\end{aligned}$$

### 2015-10-31 notes/rough-calculations.Rmd
rough calcs

>>>>

**Answer:**

a.  Maximum cline width $\sigma \sqrt{\tau}$, even at unlinked loci.
b.  The selected cline takes time $s$ to spread out to its full width.
c.  Cline behaves like the selected cline for $r < 1/\tau$ (more tentatively, for $r < \sqrt{\frac{2}{s\tau}}$).
d.  Cline is flattening for $x<\sigma\sqrt{\tau}$ and $r > \sqrt{\frac{2}{s\tau}}$.

*Rationale:*

a.  The maximum of Brownian motion until time $t$ has mean $\sqrt{2t/\pi}$,
    and has the same marginal distribution as the absolute value of Brownian motion.
    Therefore, the probability a lineage from $x$ hasn't got to the center of the cline *at all* by time $\tau$ 
    is about $2\Phi(x/\sqrt{\sigma^2 \tau})-1$;
    and therefore, if $x/\sigma > 2\sqrt{\tau}$ then the cline cannot have started to form, regardless of $r$.

b.  The selected cline spreads out (in forwards time!) with diffusion,
    and so will reach maximum width $\sigma/\sqrt{s}$ in time $s$.
    Let $q(t,x)$ be the solution to
    $$\begin{aligned}
    \dot q = \frac{1}{2} q'' + q (1-q) (2q-1) .
    \end{aligned}$$
    If $u=at$ and $y=bx$ and let $p(t,x) = q(at,bx)$ then
    $$\begin{aligned}
    \dot p(t,x)
        &= \partial_t q(at,bx) \\
        &= a \dot q(at,bx) \\
        &= a \left\{ \frac{1}{2} q''(at,bx) + q (1-q) (2q-1) \right\} \\
        &= a \left\{ \frac{1}{2b^2} \partial^2_x q(at,bx) + q (1-q) (2q-1) \right\} \\
        &= \frac{a}{2b^2} p''(t,x) + a p (1-p) (2p-1)  .
    \end{aligned}$$
    If we choose $a=s$ and $b=\sigma/\sqrt{s}$ then $a/b^2=\sigma^2$, so
    $$\begin{aligned}
    \dot p = \frac{\sigma^2}{2} p'' + s p (1-p) (2p-1) .
    \end{aligned}$$
    We know that the unitless cline $q$ will spread out in time 1;
    so our cline will spread out in time $s$.

c.  The cline will be narrower if $r$ is small.
    Consider a lineage begun near the center of the cline.
    The most time it can spend in heterozygotes is $\tau$,
    so if $r \ll 1/\tau$, the chance of recombining is small.
    So, the cline will look like the cline at the selected locus if $r < 1/\tau$.

d.  By L\'evy's identity, the local time of a Brownian motion at zero
    has the same distribution (pathwise!) as the running maximum of Brownian motion.
    Therefore, the local time at zero up until $\tau$ grows like $\sqrt{\tau}$.
    The same is true for Brownian motion reflected at zero.
    The time a lineage spends in heterozygotes should be roughly proportional to the local time
    (each time it visits zero from the left, it spends some time in hets before leaving again; 
    the constant of proportionality is the mean time spent per visit).
    Therefore, there will be appreciable chance of recombination if $r C \sqrt{\tau}>1$,
    where $C$ is the constant;
    this says that if clines down to $r$ are flat at $\tau$, then at time $2\tau$ they'll be flat down to $r/4$.

    But, what is the constant $C$?
    Consider the approximation of a lineage that acts like Brownian motion with a constant drift $s$ above zero.
    As noted in [this file](diffusion-local-times.html),
    the excursion measure of Brownian motion assigns mass $\sqrt{\frac{1}{2 \pi t^3}}$ to excursions of length $t$;
    we must tilt this by $e^{-ts}$ for the selection above zero;
    so that the mean rate at which this semi-pushed motion accumulates time above zero,
    with respect to local time at zero,
    is  (letting $x=\sqrt{2st}$)
    $$\begin{aligned}
    \int_0^\infty t \frac{e^{-st}}{ \sqrt{2 \pi t^3 } } dt 
    &=
    \int_0^\infty \frac{e^{-st}}{ \sqrt{2 \pi t } } dt \\
    &=
    \sqrt{2s} \int_0^\infty \frac{e^{-x^2/2}}{ \sqrt{2 \pi} } dx \\
    &=
    \sqrt{\frac{s}{2}} .
    \end{aligned}$$
    This suggests taking $C=\sqrt{s/2}$.
